>

### 2025-01-31 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
cleaning up

abstract: > 
  Learning language requires learning not only the content of language, but also how to use language to communicate. Iterated reference games provide a window into such skills, requiring rich communication as participants converge on mutually understandable names for initially novel referents. Some early experiments are interpreted as evidence that 4-5-year old children cannot converge to the mutually understandable names needed to communicate in an iterated reference game. Here, we revisit young children's referential communicative abilities using a simpler, child-friendly paradigm. Across 51 pairs of children, we found that 4-5 year olds successfully established reference with each other. Children were 85% accurate, and they often used descriptions similar to their partner's. These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with the view that children show early pragmatic competence in supportive contexts.

@@

One influential early study suggested that 4-5-year old preschoolers struggle with child-child referential communication [@glucksberg1966]. In their paradigm, one child was given a set of 6 blocks in a specific order. Their task was to describe the images on the block so their partner could pick out the corresponding block. As children described and selected blocks, they stacked them on pegs. While 4-5-year old children succeeded on practice trials with familiar shapes and visual access to each other's blocks, children failed on critical trials where the blocks had abstract drawings and there was no visual access. Even after multiple rounds with the same images, children were not able to correctly order the blocks. @glucksberg1966 attributed children's communicative failures to their production of ego-centric descriptions that did not account for the other child's perspective.

@@

More recently, though no evidence has directly contradicted this early study, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication. Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6 year olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication. Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009]. By age 5, children integrate information about other's perspectives into their comprehension of utterances [@sanjuan2015]. Overall, by late preschool, children can can reason about others’ perspectives to communicate more effectively.

While children are sensitive to others’ perspectives, they still struggle to appropriately tailor the specificity of their utterances to the visual context, often resulting under-informative utterances [@matthews2012; @leung2024].  When children must describe one of two very similar images, 4 and 5-year olds sometimes neglect to mention the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019]. By 5-years old, children are sensitive to over- and under-informative utterances, asking for clarification on some under-informative utterances and taking longer to make selections on over-informative utterances [@morisseau2013].

A common thread among many developmental studies is that children perform better when the cognitive load of the task is reduced, i.e. when there are fewer possible referents [@abbot-smith2016; @sanjuan2015].  These findings paired with evidence of emerging communicative skills in young children suggests that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.

Since @glucksberg1966, few studies have revisited the question of whether children can successfully communicate in an iterated reference game. In one study, 8-10-year olds exhibited adult-like patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions, but children’s accuracy was still far below adult performance  and highly variable between dyads [@branigan2016]. 4-6-year olds successfully used conventionalized gestures with their partners in an iterated reference game where children could only use gestures to communicate [@bohn2019]. 4-8 year old children succeeded at playing an iterated reference game with a parent using a simple, child-friendly tablet-based task [@leung2024]. In this task, even 4-year olds had an initial accuracy above 80%, which rose to above 90% in later repetitions [@leung2024]. Together, these findings provide evidence of young children’s ability to communicate about novel referents under the right conditions.

Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, here we revisit the question of child-child referential communication. In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm. Across an initial study and a recplication including a total of 51 pairs of 4-5-year old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.

@@

Our goal for Experiment 1 was to test young children's ability to coordinate on descriptions to abstract shapes that their partner could understand. Young children can be very sensitive to task demands and cognitive load [@turan-kucuk2024; @keen2003; @carruthers2013], so we adapted the experimental framework from @leung2024, and further simplified it by reducing the total pool of targets and the number of trials. This experiment was pre-registered at https://osf.io/kcv8j.

@@

4 and 5-year old children were recruited from a university preschool during the school day. Children played with another child from the same class. Experiment 1 was conducted between June and August 2023. Pairs of children were included in analyses if they completed at least 8 of the 12 critical trials. We had 19 games that completed 12 critical trials, and 1 game that completed 11 critical trials. Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months.

### Materials
For the target stimuli, we used four of the ten tangram images from @leung2024, chosen based on visual dissimilarity (Figure \ref{game}B). We coded the matching game using Empirica [@almaatouq2020], hosted it on a server, and then accessed the game on ablets that were locked in a kiosk mode so children could not navigate away from the game.

@@

### Participants
Experiment 2 was run between March and August of 2024, at the same preschool as Experiment 1. No children participated in both experiments. 30 pairs of children completed all 16 critical trials, and 1 pair of children completed 10 critical trials. Our target age range was 4 and 5 year olds, but one older 3-year-old was unintentionally included. Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months.

@@

Prominent early findings been interpreted as evidence that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966]. However, more recent developmental have found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low. <!--Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents. -->We revisited the question of preschoolers's performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.

Across 2 experiments and 51 pairs of 4 and 5-year old children, we tested how well children could produce referential expressions that allowed their partner to find a matching abstract shape. Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (85%), indicating that children were generally able to produce adequate descriptions. Additionally, children's utterances showed signs of converging toward conceptual pacts. While this task is substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.

>>>>

abstract: > 
  Learning language requires learning not only the content of language, but also how to use language to communicate. Iterated reference games provide a window into such skills, requiring rich communication as participants converge on mutually understandable names for initially novel referents. Some early experiments are interpreted as evidence that 4-5-year-old children cannot converge to the mutually understandable names needed to communicate in an iterated reference game. Here, we revisit young children's referential communicative abilities using a simpler, child-friendly paradigm. Across 51 pairs of children, we found that 4-5-year-olds successfully established reference with each other. Children were 85% accurate, and they often used descriptions similar to their partner's. These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with the view that children show early pragmatic competence in supportive contexts.

@@

One influential early study suggested that 4-5-year-old preschoolers struggle with child-child referential communication [@glucksberg1966]. In their paradigm, one child was given a set of 6 blocks in a specific order. Their task was to describe the images on the block so their partner could pick out the corresponding block. As children described and selected blocks, they stacked them on pegs. While 4-5-year-old children succeeded on practice trials with familiar shapes and visual access to each other's blocks, children failed on critical trials where the blocks had abstract drawings and there was no visual access. Even after multiple rounds with the same images, children were not able to correctly order the blocks. @glucksberg1966 attributed children's communicative failures to their production of ego-centric descriptions that did not account for the other child's perspective.

@@

More recently, though no evidence has directly contradicted this early study, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication. Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6-year-olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication. Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009]. By age 5, children integrate information about other's perspectives into their comprehension of utterances [@sanjuan2015]. Overall, by late preschool, children can can reason about others’ perspectives to communicate more effectively.

While children are sensitive to others’ perspectives, they still struggle to appropriately tailor the specificity of their utterances to the visual context, often resulting in under-informative utterances [@matthews2012; @leung2024].  When children must describe one of two very similar images, 4 and 5-year-olds sometimes neglect to mention the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019]. By 5 years old, children are sensitive to over- and under-informative utterances, asking for clarification on some under-informative utterances and taking longer to make selections on over-informative utterances [@morisseau2013].

A common thread among many developmental studies is that children perform better when the cognitive load of the task is reduced, i.e. when there are fewer possible referents [@abbot-smith2016; @sanjuan2015].  These findings paired with evidence of emerging communicative skills in young children suggest that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.

Since @glucksberg1966, few studies have revisited the question of whether children can successfully communicate in an iterated reference game. In one study, 8-10-year-olds exhibited adult-like patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions, but children’s accuracy was still far below adult performance and highly variable between dyads [@branigan2016]. 4-6-year-olds successfully used conventionalized gestures with their partners in an iterated reference game where children could only use gestures to communicate [@bohn2019]. 4-8-year-old children succeeded at playing an iterated reference game with a parent using a simple, child-friendly tablet-based task [@leung2024]. In this task, even 4-year-olds had an initial accuracy above 80%, which rose to above 90% in later repetitions [@leung2024]. Together, these findings provide evidence of young children’s ability to communicate about novel referents under the right conditions.

Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, here we revisit the question of child-child referential communication. In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm. Across an initial study and a replication including a total of 51 pairs of 4-5-year-old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.

@@

Our goal for Experiment 1 was to test young children's ability to coordinate on descriptions of abstract shapes that their partner could understand. Young children can be very sensitive to task demands and cognitive load [@turan-kucuk2024; @keen2003; @carruthers2013], so we adapted the experimental framework from @leung2024, and further simplified it by reducing the total pool of targets and the number of trials. This experiment was pre-registered at https://osf.io/kcv8j.

@@

4 and 5-year-old children were recruited from a university preschool during the school day. Children played with another child from the same class. Experiment 1 was conducted between June and August 2023. Pairs of children were included in analyses if they completed at least 8 of the 12 critical trials. We had 19 games that completed 12 critical trials, and 1 game that completed 11 critical trials. Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months.

### Materials
For the target stimuli, we used four of the ten tangram images from @leung2024, chosen based on visual dissimilarity (Figure \ref{game}B). We coded the matching game using Empirica [@almaatouq2020], hosted it on a server, and then accessed the game on tablets that were locked in a kiosk mode so children could not navigate away from the game.

@@

### Participants
Experiment 2 was run between March and August of 2024, at the same preschool as Experiment 1. No children participated in both experiments. 30 pairs of children completed all 16 critical trials, and 1 pair of children completed 10 critical trials. Our target age range was 4 and 5-year-olds, but one older 3-year-old was unintentionally included. Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months.

@@

Prominent early findings have been interpreted as evidence that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966]. However, more recent developmental work has found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low. <!--Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents. -->We revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.

Across 2 experiments and 51 pairs of 4 and 5-year-old children, we tested how well children could produce referential expressions that allowed their partner to find a matching abstract shape. Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (85%), indicating that children were generally able to produce adequate descriptions. Additionally, children's utterances showed signs of converging toward conceptual pacts. While this task is substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
